Diffusive Homeostasis in a Self-Organizing Recurrent Neural Network \\ \begin{large} Spatially dependent Interaction as a Determinant of Neural Activity and Plasticity \end{large}

### 2016-12-07 draft.tex
.

Balanced networks of inhibitory and excitatory populations of spiking LIF-neurons are known to show highly random, uncorrelated spiking behaviour under sparse connectivity with a relatively high per-synapse connection strength, see e.g. \cite{Vreeswijk1996} or \cite{Brunel2000}. This kind of asynchronous regular state is characterized by a population firing rate showing no regularities over time, though maintaining an overall stable level of activity. This feature was shown to be achieved after slowing down the rate of homeostatic adaption. On the level of single cells, interspike intervals should be random, approximately following the statistics of a poisson process.

>>>>

Balanced networks of inhibitory and excitatory populations of spiking LIF-neurons are known to show highly random, uncorrelated spiking behaviour under sparse connectivity with a relatively high per-synapse connection strength, see e.g. \cite{Vreeswijk1996} or \cite{Brunel2000}. This kind of asynchronous regular state is characterized by a population firing rate showing no regularities over time, though maintaining an overall stable level of activity. This feature was shown in the previous section after slowing down the rate of homeostatic adaption. On the level of single cells, interspike intervals should be random, approximately following the statistics of a poisson process.   
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/Spike_Stats/isi_cv_compare.png}
\caption{A/C: Distribution of interspike intervals of excitatory/inhibitory neurons for diffusive and non-diffusive homeostasis, measured over $\mathrm{1000\,s}$. Only neurons with a mean firing rate in a certain range (see legend) were included into statistics to prevent overlap of different firing rates (except for non-diff. homeostasis and excitatory neurons, all being very close to 3 Hz). Black lines are fits of exponential functions. B/D: Distribution of coefficients of variation. Each sample represents the CV of one excitatory/inhibitory neuron.}
\label{ISI_CV_compare}
\end{figure}
Figure \ref{ISI_CV_compare} shows the distribution of interspike intervals (ISIs) and the distribution of coefficient of variation ($\mathrm{CV\equiv\sigma/\mu}$), taken over the set of excitatory and inhibitory neurons. However, in order to provide comparability between the ISI distributions in the diffusive/non-diffusive case, we restricted the statistics of ISIs to neurons whose mean firing rate fall into a window with a width of 1 Hz and a certain mean (see legend in Figure \ref{ISI_CV_compare}). Apparently, an approximately exponential distribution of ISIs is preserved for diffusive homeostasis (represented by straight lines in a log-plot), which is characteristic for poisson processes \cite[p. 27]{Theor_Neur_Dayan}. For perfect poissonian spiking, the CV should give a value of 1. CVs that are smaller/larger, as seen in Figure \ref{ISI_CV_compare} B/D, indicate a tendency towards greater/less regularity. This is presumably due to the statistics of ISIs close to zero: Interestingly, while excitatory neurons have an  under-representation of short ISIs, the opposite is the case for inhibitory neurons. These types of behaviour of LIF-models have been shown to depend on the CV of input the neuron receives \cite{Ostojic2011}: A smaller signal-to-noise-ratio (larger CV) of neural input leads to a larger CV of ISIs for the output and an over-representation of small intervals, see Figure \ref{Ostojic_ISI_plot}.
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/external_plots/ostojic_isi.png}
\caption{Plot was taken from \cite{Ostojic2011}. Top row: sample trajectories of the membrane potential of a LIF-neuron (blue curve) and the corresponding probability density (yellow). From left to right, the input's CV is successively increased. Bottom row: the corresponding ISI distributions and the dominating exponential curve for large ISIs (red).}
\label{Ostojic_ISI_plot}
\end{figure}
This relation fits nicely into our set of network parameters: inhibitory neurons receive stronger recurrent inhibitory input ($\mathrm{-60\,mV}$ in total) compared to excitatory neurons ($\mathrm{-12\,mV}$). This lowers the signal-to-noise ratio for the inputs of inhibitory neurons, leading to the aforementioned statistical differences. Still, one can observe a dominant exponential component for both excitatory and inhibitory neurons.
